---
title: "NHS-R Community Book club"
subtitle: "Invisible Women, chapters 3 & 4"
author: "Zoë Turner"
date: "July 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    seal: false
    self_contained: true
    nature:
      highlightStyle: googlecode
      highlightLines: true
      highlightLanguage: ["r"]
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include = FALSE}
library(knitr)
library(tidyverse)

### Packages not on CRAN ###
library(xaringanthemer)

# install.packages("remotes")
# remotes::install_github("mitchelloharawild/icon")
library(icon) # icons need to be loaded to view so run following code if not already done
# icon::download_fontawesome()

# install.packages("devtools")
#devtools::install_github("gadenbuie/xaringanExtra")
library(xaringanExtra)
xaringanExtra::use_share_again() # need to get the slide button on html view

# set default options
opts_chunk$set(echo = FALSE,
               fig.width = 7.252,
               fig.height = 4,
               dpi = 300,
               dev.args = list(type = "cairo"),
               eval = TRUE,
               warning = FALSE,
               message = FALSE,
               error = FALSE)

```

```{r echo=FALSE}
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)

```

class: inverse, middle, center

# `r rmarkdown::metadata$title`
----
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author` | `r rmarkdown::metadata$date`
</br>

---

# Note following meeting

I wrote these notes for the group but didn't show them as we discussed the book generally and a few of the stories that are featured. 

These slides are built using {xaringan}, {xaringanExtra} and {xaringanthemer} - there appears to be an issue with the `--` for some slides where this is printing and text isn't sliding into view as intended.

---

# Chapter 3 - money 

Opens with a story about the women of Iceland going on strike on 24 October 1975

--

* 90% (20k+) of women went on strike

--

* A year later Iceland passed the first Gender Equality Act

--

* By 2017 Iceland topped the World Economic Forum's Global Gender Gap Index for the 8th year running

--

## That's a huge impact

---

# But things are changing...

A common retort that men are doing more of their share of the household work burden

--

* Individually, yes

* At a population level, no

-- 

Women take on a greater share of housework, caring responsibilities and also, often, work

--

Leisure pursuits are imbalanced with women's being shorter, more fragmented and often 
combine tasks

--

When ill or recovering from surgery women return to the housework sooner

--

To the extent that single female recover better from heart attacks than married women

---

# Unpaid work

The hidden data for women isn't always 'direct'

--

Indirectly women are not being counted in unpaid work they do when "science" says you shouldn't "work" more than 40 (paid of course) hours a week 

--

To accommodate unpaid work responsibilities women work part-time (75% are women)

--

Part-time work is often lower graded so women get paid less

--

That gets hidden in pay gap statistics when part-time and full-time is combined

---

# Money

Lower pension pots and savings due to working patterns

--

Leads to greater poverty in older years

--

Remember women live longer then men (on average)

--

And as for maternity it pays poorly to have children 

* UK was 22nd out of 24 European countries on length of "decently paid maternity leave"

-- 

For some it doesn't pay at all, like MPs and councillors

--

Paternity schemes where time is shared often work better when rules where its not taken up by the other partner then it's lost

-- 

Huge effects on women working in universities as the system of tenure doesn't allow for gaps to have children

-- 

Some processes put in place that are meant to be fair by including the father, further disadvantaged the mother - they were able to use the time to publish whereas the mothers didn't

---

# Childcare

It's not just about babies! 

--

Taxes disadvantage women who need childcare as it's a personal expense

-- 

But booking a hotel for a night is a valid expense

---

# Chapter 4 - meritocracy 

Blind auditions for the orchestra equalised numbers of man to women

--

Desire for meritocracy to exist, to work very strong

--

Criticism in performance reviews is more personal for women - tone & character

--

Believing in meritocracy can introduce bias (justification of decisions?)

-- 

Differences in publishing and citations (men self cite 70% more than women)

--

Male default thinking so when using initials, assumed male

---

# Time

Differences in time requested to do papers (requesting extensions, re-grading and so on)

-- 

Admin burden falls to women - extra teaching hours, paperwork

-- 

Emotional support expected from staff but then also penalised

--

Least effective males are still ranked higher than more effective females


---

# Teaching

We still use teaching evaluations and they are useless (resonates)

--

Personal comments appear more frequently in female lecturers feedback

--

Girls are learning that brilliance and genius is female at school

--

Females were the original computers and coders. Great book recommendation is [Coders](https://www.panmacmillan.com/authors/clive-thompson/coders/9781529018981) by Clive Thompson

---

# Thoughts

Highlights poor logical thinking - so much of what is highlighted in this book is that assumptions are made so often on what we see and record, forgetting the missing

--

Lots of facts

--

Stories are more memorable

--

After a week, what do you remember?